``Interestingness'' in Mathematical Theory Exploration

### 2019-10-18 PhDThesis/litreview_staticanalysis.tex
Remove fixed TODO
Add citations

The most common code-improvement technique performed \emph{during execution} is
Just In Time (JIT) compilation~\iffalse TODO: Cite \fi, which allows an
interpreter to compile sections of code into a faster form (usually machine
code). Since this compilation takes time, it is usually reserved for sections
which are executed often enough to exceed some predefined threshold (e.g. 100).
By focusing on such ``hot paths'', the up-front costs of compiling everything
ahead of time are avoided; although compiled forms are kept only in memory,
getting recompiled as needed on each program run. Compiling at runtime also
allows specialisation with dynamically-known information; for example, inlining
specific implementations of a procedure rather than performing dynamic dispatch
(this is the problem higher-order functions discussed above). To avoid altering
semantics, such dynamic specialisations must be accompanied by a conservative
validity check (e.g. that the given arguments should use the procedure
implementation we have compiled-in), falling back to the interpreter upon
failure.

A more speculative use of optimisation during runtime has been proposed by
Hutter~\iffalse TODO: Cite \fi. As in a JIT-enabled interpreter, Hutter's
algorithm (also referred \hsearch{}) begins by directly executing the given
(unoptimised) program $p$ on the relevant input $x$ (this may be provided up
front, or piecemeal throughout execution, e.g. as a stream of user interaction
events). The variables $p_{fast}$ and $t_{fast}$ are initialised to $p$ and
$\infty$ respectively, to indicate that $p$ is the fastest implementation we
have so far, and we don't yet have an upper bound on its running
time. Concurrently with the execution of $p$, two other processes are also
started:

\begin{itemize}
\item Process A searches for proofs (in a pre-specified formal system) that some
  program $p\prime$ is both semantically equivalent to $p$ and has a running
  time bounded by some other program $t$, i.e. for all inputs $i$, the
  time taken to compute $p\prime(i)$ is bounded by $t(i)$. A list is built up of
  the $(p\prime, t)$ pairs found so far.
\item Process B computes $t(x)$ for all of the bounds $t$ found so far by A.
  Computations are interleaved via a schedule similar to Levin's universal
  search algorithm~\iffalse TODO: Cite \fi, to prevent getting stuck on
  non-halting programs. Whenever a calculation of some $t(x)$ finishes, it is
  compared to $t_{fast}$: if smaller, $t_{fast}$ is set to $t(x)$, and $p_{fast}$ is
  set to $p\prime$ (the program paired with this $t$).
\end{itemize}

@@

% TODO: Evolutionary patches

>>>>

The most common code-improvement technique performed \emph{during execution} is
Just In Time (JIT) compilation~\cite{aycock2003brief}, which allows an
interpreter to compile sections of code into a faster form (usually machine
code). Since this compilation takes time, it is usually reserved for sections
which are executed often enough to exceed some predefined threshold (e.g. 100).
By focusing on such ``hot paths'', the up-front costs of compiling everything
ahead of time are avoided; although compiled forms are kept only in memory,
getting recompiled as needed on each program run. Compiling at runtime also
allows specialisation with dynamically-known information; for example, inlining
specific implementations of a procedure rather than performing dynamic dispatch
(this is the problem higher-order functions discussed above). To avoid altering
semantics, such dynamic specialisations must be accompanied by a conservative
validity check (e.g. that the given arguments should use the procedure
implementation we have compiled-in), falling back to the interpreter upon
failure.

A more speculative use of optimisation during runtime has been proposed by
Hutter~\cite{HUTTER:2002}. As in a JIT-enabled interpreter, Hutter's
algorithm (which we refer to as \hsearch{}) begins by directly executing the
given (unoptimised) program $p$ on the relevant input $x$ (this may be provided
up front, or piecemeal throughout execution, e.g. as a stream of user
interaction events). The variables $p_{fast}$ and $t_{fast}$ are initialised to $p$
and $\infty$ respectively, to indicate that $p$ is the fastest implementation we
have so far, and we don't yet have an upper bound on its running
time. Concurrently with the execution of $p$, two other processes are also
started:

\begin{itemize}
\item Process A searches for proofs (in a pre-specified formal system) that some
  program $p\prime$ is both semantically equivalent to $p$ and has a running
  time bounded by some other program $t$, i.e. for all inputs $i$, the
  time taken to compute $p\prime(i)$ is bounded by $t(i)$. A list is built up of
  the $(p\prime, t)$ pairs found so far.
\item Process B computes $t(x)$ for all of the bounds $t$ found so far by A.
  Computations are interleaved via a schedule similar to Levin's universal
  search algorithm~\cite{levin1984randomness,ming}, to prevent getting stuck on
  non-halting programs. Whenever a calculation of some $t(x)$ finishes, it is
  compared to $t_{fast}$: if smaller, $t_{fast}$ is set to $t(x)$, and $p_{fast}$ is
  set to $p\prime$ (the program paired with this $t$).
\end{itemize}

@@
